Experiment 2 of 4

### 2016-10-02 experiments/e2/writeup/e2-analysis.Rmd
e2: simpler file structure; now includes coefficient plots for main analyses; raw data is in writeup

---
title: "Experiment 2 of 4"
author: "Matt Green"
output:
  html_document:
    toc: yes
    <!--toc_float: true-->
    mathjax: default
    template: default
    <!--code_folding: show-->
    theme: default
---  
<style>
table, th, td {
    border: 1px grey;
    border-collapse: collapse;
    padding:2px 5px 2px 5px
}  
</style>

```{r setup, include=FALSE, cache=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(pander)
```

```{r showInstructions, echo=FALSE, cache=TRUE}
v = c("crisp", "crisp", "vague", "vague")
n = c('number', 'word', 'number', 'word')
e <- c("Choose the square with 6 dots", "Choose the square with the fewest dots", "Choose a square with about 10 dots", "Choose a square with few dots")
i = data.frame(v, n, e)
names(i) = c("vagueness", "number", "example instruction")
pander(i, justify=c('center', 'left', 'left'))
```

```{r showItems, echo=FALSE, cache=TRUE}
items = c(1,2,3,4,5,6,7,8)
i = data.frame(Item = c("1","2","3","4"),
               x = c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
names(i) = c('Item', 'Numbers of dots')
pander(i, justify=c('center', 'center'))
```

```{r loadLibraries, message=FALSE, echo=FALSE, cache=FALSE}
library(Matrix); library(knitr); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); source('summarySEwithin2.R'); source('e2-preprocessing.R')
```

```{r preprocessing}
source('e2-preprocessing.R')
if( !file.exists('e2-data.txt') ) {
  dat <- gatherData() 
  dat <- declareImpossibleRT(dat)
  #dat <- removeImpossibleTrials(dat)
  dat <- subset(dat, select=c(Subject, Trial, Item, c_Itm, 
                              discriminability, 
                              Instruction,
                              nchar_instr,
                              Vagueness, Number, Order, Quantity, 
                              c_Vag, c_Num, c_Ord, c_Qty,
                              NaType, response_category, 
                              isBorderline, RT, RT_log)
  )
  write.table(dat, file='e2-data.txt', sep='\t', quote=FALSE, row.names = FALSE)
}
```

@@

```{r getData, echo=T}
rawdata <- read.delim('e2-b-data.txt')
tempdata <- perSubjectTrim.fnc(rawdata[complete.cases(rawdata),], response='RT', subject='Subject', trim = 2.5)$data
rtdata <- tempdata
bldata <- tempdata
```

This code creates the condition means with confidence intervals used for plotting:

```{r computeSummariesForGraphs, cache=TRUE, message=FALSE}
rtdataplot1 <- summarySEwithin2(rtdata, 
                                measurevar="RT_log", 
                                withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
```

```{r rtplot, echo=F}
dodge = position_dodge(width=0.2)
rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
rtplot=ggplot(rtdataplot1, aes(y=RT_log, x=Item, ymin=RT_log-ci, ymax=RT_log+ci,
                group=Condition, fill=Vagueness)) +
  geom_line(position=dodge) +
  geom_errorbar(width=.2, position=dodge) +
  geom_point(pch=21, size=3, position=dodge) +
  #scale_shape_manual("",values = c(22, 22)) +
  scale_fill_manual("",values=c("black","white")) +
  ggtitle("Response time") +
  ylab("Respone time (log (ms))") + 
  xlab("") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), legend.position="top",
        aspect.ratio=1,
        #axis.text.x = element_text(angle = 45, hjust = 1),
        plot.background=element_rect(fill=NA, color='white')) +
  facet_grid(~Number)
rtplot
```

```{r blplot, fig.width=5, fig.height=2.5, echo=F}
bldata$Condition <- as.factor(paste(sep=':', bldata$Vagueness, bldata$Number))
bldata$response_category <- relevel(bldata$response_category, ref = 'expected')
blplot=ggplot(bldata) + 
  geom_bar(aes(Vagueness, fill=response_category), position = 'dodge') +
  #geom_errorbar() +
  facet_wrap(nrow=1,~Number) +
  scale_fill_grey(name="") +
  xlab(" ") +
  ggtitle("Response Type") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), legend.position="top", aspect.ratio=1,
        plot.background=element_rect(fill=NA, color='white'))
blplot
```

```{r rtlmer, cache=TRUE}
rtlmer = lmerTest::lmer(data=rtdata,
                        RT_log ~ c_Vag * c_Num + c_Itm +
                          (1 + c_Vag * c_Num + c_Itm | Subject))
```

```{r rtlmersummaryxtable, results='asis', cache=TRUE}
print(xtable(summary(rtlmer)$coef, digits=c(0,2,2,1,1,4)),type='html')
```

```{r seplmersfornum, results='asis', cache=TRUE}
s1 <- subset(rtdata, subset=Number=='Numeric')
s1lm <- lmerTest::lmer(data=s1, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
print(xtable(summary(s1lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
```

```{r seplmersforvag, results='asis', cache=TRUE}
s2 <- subset(rtdata, subset=Number=='Verbal')
s2lm <- lmerTest::lmer(data=s2, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
print(xtable(summary(s2lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
```

```{r bllmer, cache=TRUE}
bllmer = lme4::glmer(data=bldata,
                        isBorderline ~ c_Vag * c_Num + c_Itm +
                          (1 + c_Vag * c_Num + c_Itm | Subject),
                        family="binomial", control = glmerControl(optimizer = "bobyqa"))
```

```{r bllmersummaryxtable, results='asis', cache=TRUE}
print(xtable(summary(bllmer)$coef, digits=c(0,2,2,1,4)),type='html')
```

>>>>

---
title: "Experiment 2 of 4"
author: "Matt Green"
output:
  html_document:
    toc: yes
    <!--toc_float: true-->
    mathjax: default
    template: default
    <!--code_folding: show-->
    theme: default
---

<style>
table, th, td {
    border: 1px grey;
    border-collapse: collapse;
    padding:2px 5px 2px 5px
}  
</style>

@@

```{r loadLibraries, message=FALSE, echo=FALSE}
library(knitr); library(Matrix); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); source('summarySEwithin2.R'); source('e2-preprocessing.R')
```

```{r showInstructions, echo=FALSE}
v = c("crisp", "crisp", "vague", "vague")
n = c('number', 'word', 'number', 'word')
e <- c("Choose the square with 6 dots", "Choose the square with the fewest dots", "Choose a square with about 10 dots", "Choose a square with few dots")
i = data.frame(v, n, e)
names(i) = c("vagueness", "number", "example instruction")
pander(i, justify=c('center', 'left', 'left'))
```

```{r showItems, echo=FALSE}
items = c(1,2,3,4,5,6,7,8)
i = data.frame(Item = c("1","2","3","4"),
               x = c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
names(i) = c('Item', 'Numbers of dots')
pander(i, justify=c('center', 'center'))
```

```{r preprocessing, echo=FALSE, message=FALSE}
source('e2-preprocessing.R')
if( !file.exists('e2-data.txt') ) {
  dat <- gatherData() 
  dat <- declareImpossibleRT(dat)
  #dat <- removeImpossibleTrials(dat)
  dat <- subset(dat, select=c(Subject, Trial, Item, c_Itm, 
                              discriminability, 
                              Instruction,
                              nchar_instr,
                              Vagueness, Number, Order, Quantity, 
                              c_Vag, c_Num, c_Ord, c_Qty,
                              NaType, response_category, 
                              isBorderline, RT, RT_log)
  )
  write.table(dat, file='e2-data.txt', sep='\t', quote=FALSE, row.names = FALSE)
}
```

```{r getData, echo=TRUE, tidy=FALSE}
rawdata <- read.delim('e2-data.txt')
tempdata <- perSubjectTrim.fnc(data = rawdata[complete.cases(rawdata),], 
                               response='RT', subject='Subject', trim = 2.5)$data
rtdata <- tempdata
bldata <- tempdata
```

```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE, dev='pdf'}
rtdataplot1 <- summarySEwithin2(rtdata, 
                                measurevar="RT_log", 
                                withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, group=Condition, fill=Vagueness)) +
  geom_line() +
  geom_errorbar(width=.2, aes(ymin=RT_logNormed-ci, ymax=RT_logNormed+ci)) +
  geom_point(size=2, col=1, aes(shape=Vagueness)) +
  scale_fill_grey(name="Vagueness", start=0, end=1) +
  scale_shape_manual(name="Vagueness", values=c(21,22)) +
  ggtitle("Response time") +
  ylab("Respone time (log (ms))") + 
  xlab(NULL) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), legend.position="right",
        aspect.ratio=1,
        plot.background=element_rect(fill=NA, color='white'),
        strip.background=element_blank(),
        axis.text.x = element_text(angle = 15)) +
  facet_grid(~Number)
rtplot
```

```{r e2-blBarChart, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE, dev='pdf'}
bldata$response_category <- relevel(bldata$response_category, ref='expected')
blBarChart = ggplot(bldata) +
  geom_bar(aes(response_category, group=Number:Vagueness, fill=Vagueness),
           position=position_dodge(width=NULL)) +
  scale_fill_grey() +
  xlab(NULL) +
  ggtitle('Borderline response distribution') +
  facet_grid(~Number) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), legend.position="right", 
        aspect.ratio=1,
        plot.background=element_rect(fill=NA, color='white'),
        strip.background=element_blank(),
        legend.key.size=unit(4, 'mm'),
        axis.text.x = element_text(angle = 15))
blBarChart
```

Response time models

```{r e2-rtlmerTest1, cache=TRUE, echo=TRUE}
rtlmerTest1 = lmerTest::lmer(
  data=rtdata,
  RT_log ~ c_Vag * c_Num + c_Itm +
    (1 + c_Vag * c_Num + c_Itm | Subject))
```

```{r e2-rtlmerTest1Visuals, cache=TRUE, echo=FALSE, results='asis', dependson='e2-rtlmerTest1', fig.width=6, fig.height=3, dev='pdf'}
print(
  xtable(
    summary(rtlmerTest1)$coefficients[,-3], 
    display=c("s","f","f","f","f"),
    digits=c(0,2,2,1,3),
    caption="Response time model"),
  caption.placement="top", 
  type='html')
dfr <- data.frame(coef = summary(rtlmerTest1)$coef[,1],
                  ci2.5 = confint(rtlmerTest1, method='Wald')[,1][!is.na(confint(rtlmerTest1, method='Wald')[,1])],
                  ci97.5 = confint(rtlmerTest1, method='Wald')[,2][!is.na(confint(rtlmerTest1, method='Wald')[,2])]
)
intercept = round(dfr[1,1],2)
dfr <- dfr[-1,] # lose the intercept for plotting
dfr <- dfr[rev(rownames(dfr)),]
par(mar=c(5,7,2,1))
plot(y=1:nrow(dfr), 
     x=dfr$coef, 
     xlim=extendrange(range(dfr$ci2.5, dfr$ci97.5)), 
     ylim=extendrange(c(1,nrow(dfr))),
     type='n', axes=F, 
     xlab=substitute(paste('deviation from an intercept of ', intercept, ' log RT, with 95% CI.'), list(intercept=intercept)),
     ylab="",
     main = 'Response time model')
abline(v=0, lty=1, col='lightgrey')
arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
       x0=dfr$ci2.5, x1=dfr$ci97.5, 
       length=0.035,
       col='black',
       lwd=2, code=3, angle=90)
points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
axis(1)
box(which='plot', col='grey')
```

```{r e2-seplmerforNum, results='asis', cache=TRUE}
numOnly <- subset(rtdata, subset=Number=='Numeric')
numOnlylm <- lmerTest::lmer(data=numOnly, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
print(
  xtable(
    summary(numOnlylm)$coefficients[,-3], 
    display=c("s","f","f","f","f"),
    digits=c(0,2,2,1,3),
    caption="Response time model: Numeric only"),
  caption.placement="top", 
  type='html')
```

```{r e2-seplmerforVerbal, results='asis', cache=TRUE}
verbalOnly <- subset(rtdata, subset=Number=='Verbal')
verbalOnlylm <- lmerTest::lmer(data=verbalOnly, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
print(
  xtable(
    summary(verbalOnlylm)$coefficients[,-3], 
    display=c("s","f","f","f","f"),
    digits=c(0,2,2,1,3),
    caption="Response time model: Verbal only"),
  caption.placement="top", 
  type='html')
```

Borderline response models

```{r e2-bllmer, cache=TRUE, echo=TRUE}
bllmer = lme4::glmer(data=bldata,
                        isBorderline ~ c_Vag * c_Num + c_Itm +
                          (1 + c_Vag * c_Num + c_Itm | Subject),
                        family="binomial", control = glmerControl(optimizer = "bobyqa"))
```

```{r e2-bllmerVisuals, cache=TRUE, echo=FALSE, results='asis', dependson='e2-bllmer', fig.width=6, fig.height=3, dev='pdf'}
print(
  xtable(
    summary(bllmer)$coefficients,
    digits=c(0, 2, 2, 1, 3),
    caption="Error rates model"),
  caption.placement="top", 
  type='html')
dfr <- data.frame(coef = summary(bllmer)$coef[,1],
                  ci2.5 = confint(bllmer, method='Wald')[,1][!is.na(confint(bllmer, method='Wald')[,1])],
                  ci97.5 = confint(bllmer, method='Wald')[,2][!is.na(confint(bllmer, method='Wald')[,2])]
)
intercept = round(dfr[1,1],2)
dfr <- dfr[-1,] # lose the intercept for plotting
dfr <- dfr[rev(rownames(dfr)),]
par(mar=c(5,7,2,1))
plot(y=1:nrow(dfr), 
     x=dfr$coef, 
     xlim=extendrange(range(dfr$ci2.5, dfr$ci97.5)), 
     ylim=extendrange(c(1,nrow(dfr))),
     type='n', axes=F, 
     xlab=substitute(paste('deviation from an intercept of ', intercept, ' with 95% CI.'), list(intercept=intercept)),
     ylab="",
     main='Borderline response model')
abline(v=0, lty=1, col='grey')
arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
       x0=dfr$ci2.5, x1=dfr$ci97.5, 
       length=0.035,
       col='black',
       lwd=2, code=3, angle=90)
points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
axis(1)
box(which='plot', col='grey')
```

### 2016-10-02 experiments/e2/writeup/e2-scratch.Rmd
e2: simpler file structure; now includes coefficient plots for main analyses; raw data is in writeup

<!--html_document:-->
  <!--toc: true-->
  <!--toc_float: true-->
  <!--mathjax: default-->
  <!--template: default-->
  <!--code_folding: hide-->
  <!--theme: default-->

@@

<<exampleOfFonts, fig.width=3, fig.height=3, fig.cap='font example', echo=FALSE, eval=FALSE>>=
p <- ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() +
ggtitle("Fuel Efficiency of 32 Cars") +
xlab("Weight (x1000 lb)") + ylab("Miles per Gallon") +
theme_bw() +
theme(text=element_text(family="CMU Serif", size=14, face='plain'))
p
ggsave("ggplot_CMU_Serif.pdf", p, width=3.5, height=3.5)
@

<<defineFonts>>=
# .define.fonts <- function () {
#     quartzFonts(avenir =    c('Avenir Book', 'Avenir Black', 'Avenir Book Oblique', 'Avenir Black Oblique'),
#                 helvetica = c('Helvetica New Light', 'Helvetica New Bold', 'Helvetica New Light Italic', 'Helvetica New Bold Italic'),
#                 cmodern = c('CMU Serif', 'CMU Serif Extra','CMU Classical Serif','CMU Serif Extra'))
# }
# .define.fonts()
@

<<fig.cap='base graphic special font'>>=
#base graphic special font
pdf("plot_cm.pdf", family="CM Roman", width=5.5, height=5)

curve(dnorm, from=-3, to=3, main="Normal Distribution")
text(x=0, y=0.1, cex=1.5, expression(italic(y == frac(1, sqrt(2 * pi)) *
e ^ {-frac(x^2, 2)} )))

dev.off()
embed_fonts("plot_cm.pdf", outfile="plot_cm_embed.pdf")
@

>>>>

```{r setup, include=FALSE, cache=FALSE}
library(knitr)
```

```{r loadFonts, message=FALSE, echo=FALSE}
library(fontcm)
library(extrafont)
#font_import()
#loadfonts()
```

```{r loadLibs, message=FALSE, echo=FALSE}
library(beepr)                     #
library(car)                       #
library(ggplot2)                   #
library(grid)                      #
library(gridExtra)                 #
library(languageR)                 #
library(lattice)                   #
library(latticeExtra)              # for doubleYScale
library(lme4)                      #
library(lmerTest)                  #
library(LMERConvenienceFunctions)  # for perSubjectTrim
library(MASS)                      #
library(plyr)                      #
library(reshape2)                  #
library(rms)                       # for varclus
library(Rmisc)                     # for summarySEwithin and friends
library(xtable)                    #
```

```{r preprocessing}
source('e2-preprocessing.R')
if( !file.exists('e2-b-data.txt') ) {
  dat <- gatherData() 
  dat <- declareImpossibleRT(dat)
  dat <- removeImpossibleTrials(dat)
  dat <- subset(dat, select=c(Subject, Trial, Item, 
                              discriminability, 
                              Instruction,
                              nchar_instr,
                              Vagueness, Number, Order, Quantity, 
                              c_Vag, c_Num, c_Ord, c_Qty,
                              NaType, response_category, 
                              isBorderline, RT)
  )
  write.table(dat, file='e2-b-data.txt', sep='\t', quote=FALSE)
}
dd <- read.delim('e2-b-data.txt')
```

```{r showCompression, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='Ratios for different numbers of dots in the arrays: smaller values are more discriminable. Blue is for the ratio between the smallest number in the array and the largest number in an array. Red is for the mean of two ratios, one for the smallest number to the middle number, the other for the middle number to the largest number in the array', echo=FALSE}
#  Show compression of ratios of numbers of dots in the items
discriminability = sort(unique(as.numeric(dd$discriminability)))
plot(y=c(1.5,1.5,1.5,1.5), 
     x=discriminability, 
     ylab='',
     xlab='',
     yaxt='n',
     type='n', 
     xaxt='n',
     axes=F)
lines(y=c(1.5,1.5,1.5,1.5), x=discriminability, type='o', pch=19)
text(y=c(1.15,1.15,1.15,1.15), x=discriminability, labels=round(discriminability,2),  cex=.7)
mtext('mean (ratio min/med,\nratio med/max)', side = 2, line = 1)
mtext('Discriminability index', side=1, line=3)
mtext('Discriminability for each item', side=3, line=1, cex=1.5)
axis(1)
box(which='plot', col='grey')
```

```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE, dev='pdf'}
dd$RT_log=log(dd$RT)
rtdata=dd
source('summarySEwithin2.R')
rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", 
                                withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, group=Condition, fill=Vagueness)) +
  geom_line() +
  geom_errorbar(width=.2, aes(ymin=RT_log-ci, ymax=RT_log+ci)) +
  geom_point(size=2, col=1, aes(shape=Vagueness)) +
  scale_fill_grey(name="Vagueness", start=0, end=1) +
  scale_shape_manual(name="Vagueness", values=c(21,22)) +
  ggtitle("Response time") +
  ylab("Respone time (log (ms))") + 
  xlab(NULL) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), legend.position="top",
        aspect.ratio=1,
        plot.background=element_rect(fill=NA, color='white'),
        strip.background=element_blank(),
        axis.text.x = element_text(angle = 15)) +
  facet_grid(~Number)
rtplot
```
#Lmer model: before outlier removal

```{r}
dd$RT_log <- log(dd$RT)
dd$c_Trl <- as.numeric(scale(dd$Trial))
```

```{r lmer5, tidy=FALSE, cache=TRUE}
vlmerTest <- lmerTest::lmer(data=dd,
                 RT_log ~
                   c_Vag + c_Num + c_Qty + c_Ord +
                   c_Num:c_Vag:c_Qty +
                   discriminability +
                   c_Trl +
                   #RTprev_log +
                   nchar_instr +
                   (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
```

```{r printXtableV5, echo=FALSE, results='asis'}
print(
  xtable(
    summary(vlmerTest)$coef,
    caption = 'lmerTest version',
    digits=c(0, 3,3,1,1,4)),
  type='html',
  caption.placement='top'
)
```

```{r llmer5, tidy=FALSE, cache=TRUE}
vlme4 <- lme4::lmer(data=dd,
                 RT_log ~
                   c_Vag + c_Num + c_Qty + c_Ord +
                   c_Num:c_Vag:c_Qty +
                   discriminability +
                   c_Trl +
                   #RTprev_log +
                   nchar_instr +
                   (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
```

```{r printXtableV521c, echo=FALSE, results='asis'}
print(
  xtable(
    summary(vlme4)$coef,
    caption = 'lme4 version',
    digits=c(0, 3,3,1)),
  type='html',
  caption.placement='top'
)
```

@@

#FONTS

```{r defineFonts}
.define.fonts <- function () {
quartzFonts(avenir =    c('Avenir Book', 'Avenir Black', 'Avenir Book Oblique', 'Avenir Black Oblique'),
helvetica = c('Helvetica New Light', 'Helvetica New Bold', 'Helvetica New Light Italic', 'Helvetica New Bold Italic'),
cmodern = c('CMU Serif', 'CMU Serif Extra','CMU Classical Serif','CMU Serif Extra'))
}
# .define.fonts()
```

```{r exampleOfGGplotFonts, fig.width=3, fig.height=3, fig.cap='ggplot special font example', echo=FALSE}
# ggplot special fonts
p <- ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() +
ggtitle("Fuel Efficiency of 32 Cars") +
xlab("Weight (x1000 lb)") + ylab("Miles per Gallon") +
theme_bw() +
theme(text=element_text(family="CMU Serif", size=14, face='plain'))
p
ggsave("ggplot_CMU_Serif.pdf", p, width=3.5, height=3.5)
```

```{r exampleOfBaseGraphicsFonts, fig.width=3, fig.height=3, fig.cap='base graphic special font example'}
# base graphic special font
pdf("plot_cm.pdf", family="CM Roman", width=5.5, height=5)
curve(dnorm, from=-3, to=3, main="Normal Distribution")
text(x=0, y=0.1, cex=1.5, expression(italic(y == frac(1, sqrt(2 * pi)) *
e ^ {-frac(x^2, 2)} )))
dev.off()
embed_fonts("plot_cm.pdf", outfile="plot_cm_embed.pdf")
```

```{r howManyCellsInDesign}
# number of higher-level cells in the design (abstracting over Order) =
# Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
# number of measurements in eaach cell = 8
# 332 * 8 = 256
# 256 * 30 = 7680
dd$measurement=0
dd$cell=0
for (s in levels(dd$Subject)) {
  cellcount=0
  for (i in levels(dd$Item)) {
    for (n in levels(dd$Number)) {
      for (v in levels(dd$Vagueness)) {
        for (q in levels(dd$Quantity)) {
          measurementcount=0
          cellcount=cellcount+1
          for (t in dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q, "Trial"]) {
            dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "cell"] <- cellcount
            measurementcount=measurementcount+1
            dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "measurement"] = measurementcount
          }
        }
      }
    }
  }
}
```

```{r rSquaredInfo, echo=FALSE}
cat("R^2")
cor(fitted(v5), dd$RT_log)^2
```

```{r plotModelCoefsAndCis, fig.width=5.5, fig.height=4, fig.cap='Coefficient estimates and their (Wald) 95 per cent confidence intervals', fig.pos='hbtp', echo=FALSE}
# Plot model coefficients and ci's
dfr <- data.frame(coef = summary(v5)$coef[-1, 1], 
                  ci2.5  = confint(v5, method='Wald')[18:26, 1],
                  ci97.5 = confint(v5, method='Wald')[18:26, 2] )
dfr <- dfr[rev(rownames(dfr)),]
par(mar=c(5,10,1,1))
plot(y=1:nrow(dfr), 
     x=dfr$coef, 
     xlim=range(dfr$ci2.5, dfr$ci97.5), 
     type='n', axes=F, xlab="", ylab="")
abline(v=0, lty=3, col='grey')
abline(h=c(1,2,3,4,5,6,7,8,9), col='grey', lty=3)
arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
       x0=dfr$ci2.5, x1=dfr$ci97.5, 
       length=0.035,
       col='black',
       lwd=2, code=3, angle=90)
points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
axis(1)
mtext("Coefficient estimates, in units of log(RT),\nwith 95% confidence intervals", side=1, line=3.5, cex.lab=1, las=1)
mtext("Model terms", side=2, line=8)
box(which='plot', col='grey')

```

```{r plotLMERfnc, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='plotMLERfnc', echo=TRUE}
par(mfrow=c(2,4))
plotLMER.fnc(v5)

```

```{r baayenPLots99, fig.width=6, fig.height=4, fig.cap='Baayen Model Criticism Plots', fig.pos='hbtp', echo=FALSE}
# Baayen 4-plot model criticism
par(mfrow=c(1,3), pty='s')
# create scaled residuals
dd$rstand = as.vector(scale(resid(v5)))
# plot scaled residuals density
plot(density(dd$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(dd$rstand, cex=.5)
qqline(dd$rstand)
# plot standardised residuals versus fitted values
plot(dd$rstand ~ fitted(v5), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))

```

#lmerTest Version

```{r lmerTestVersionOfModel, tidy=FALSE, cache=TRUE, echo=TRUE}
v6 <- lmerTest::lmer(data=dd,
                     RT_log ~
                       c_Vag + c_Num + c_Qty + c_Ord +
                       c_Num:c_Vag:c_Qty +
                       discriminability +
                       s_Trl +
                       RTprev_log +
                       nchar_instr +
                       (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
```

```{r summaryLmerTest, cache=TRUE}
summary(v6)
```

#Lmer model: after outlier removal
not done yet.

#Borderline responses

```{r barplotBorderline, fig.width=6, fig.height=4, echo=FALSE}
dd$response_category <- relevel(dd$response_category, ref = 'expected')
barplot(table(dd$Vagueness,dd$response_category), beside=TRUE, legend=TRUE, space=c(0,.5))

```

```{r tabBdl, echo=FALSE, results='asis'}
print(xtable(table(dd$response_category,dd$Vagueness),
             caption='Borderline cases counts'),
      table.placement='htbp',
      latex.environments='center',
      size='small')
```
